

%\title{Localized Region Differential Evolution Hyperparameter Optimization for Generating Uniform Projection Designs}

\title {Studies in Hyperparameter Tuning, Design Selection and Optimization}

\author{Samuel O. Onyambu}
\department     {Statistics}
\degreeyear     {2024}

\chair          {Hongquan Xu}
\member         {Qing Zhou}
\member         {Jessica Jingyi Li}
\member         {Mark S. Handcock}






%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\dedication{\textsl{To my parents, Joseph and Jane, ---\\  
          my siblings, Tabitha, Naomi, Jasper, and Lilian,---\\  
          and my dear friends, Taylor and Justin,---\\  
          who among so many other things ensured that I gave my all to accomplish this task.}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vitaitem   {2011--2015}
                {B.S.~(Statistics and Programming), Kenyatta University, Nairobi, Kenya.}
\vitaitem   {2016--2018}
                {M.S.~(Statistics), California State University, Fullerton (CSUF).}
\vitaitem   {2016--2018}
                {Teaching Assistant, Mathematics Department, CSUF.}

\vitaitem   {2018--2019}  {Dorothy Radcliffe Dee Fellowship}
 \vitaitem               {2018-2020}  {Irma Polaski Fellowship}


\vitaitem   {2019--2021}
               {Teaching Assistant, Statistics Department, UCLA.}
\vitaitem{2021 Summer} {Mentored Research Fellowship}
\vitaitem  {2021--2022}
               {Teaching Associate, Statistics Department, UCLA.} 

\vitaitem   {2022 Summer}
                {Graduate Student Researcher, Department of Statistics, UCLA.}

\vitaitem{2022--2023} 
 {Teaching Fellow, Statistics Department, UCLA.} 

\vitaitem   {2023--2024}
                {Graduate Student Researcher, Civil and Environmental Engineering Department, UCLA.}

\vitaitem   {2024 Summer}
                {Graduate Student Researcher, Department of Statistics, UCLA.}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage

\publication    {\fullcite{dahal2022quantifying}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\abstract       {(Abstract omitted for brevity)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
